---
title: "project"
output: html_document
---
This is my first try on loading data and modelling
Lets load some data

```{r}
iris <- read.csv(url("http://archive.ics.uci.edu/ml/machine-learning-databases/iris/iris.data"), header = FALSE) 
```

Analysis of  iris
```{r}
head(iris)
str(iris)
summary(iris)
dim(iris)
```

Naming columns and data exploration
```{r}
names(iris) <- c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width", "Species")

plot(iris$Petal.Length, 
     iris$Petal.Width, 
     pch=21, bg=c("red","green3","blue")[unclass(iris$Species)], 
     xlab="Petal Length", 
     ylab="Petal Width")
```

Checking for internal and overall correlation
```{r}
M <- cor(iris[,1:4])
corrplot::corrplot(M,method ="circle")
```

Nomralizing data
```{r}
normalize <- function(x) {
  num <- x - min(x)
  denom <- max(x) - min(x)
  return (num/denom)
}
iris_norm <- as.data.frame(lapply(iris[1:4], normalize))
head(iris_norm)
iris[,5] <- as.numeric(iris[,5]) -1
iris <- as.matrix(iris)
dimnames(iris) <- NULL
head(iris)
iris <- normalize(iris[,1:4])
summary(iris)
```

Training and Testing data sets
```{r}
ind<- sample(2, nrow(iris), replace=TRUE, prob=c(0.67, 0.33))
iris.training <- iris[ind==1, 1:4]
iris.test <- iris[ind==2, 1:4]
iris.trainingtarget <- iris[ind==1, 4]
iris.testtarget <- iris[ind==2, 4]
```

Fiiting and compiling models

```{r}
library(tensorflow)
library(keras)
iris.trainLabels <- to_categorical(iris.trainingtarget)
iris.testLabels <- to_categorical(iris.testtarget)
model <- keras_model_sequential() 
model %>% 
    layer_dense(units = 8, activation = 'relu', input_shape = c(4)) %>% 
    layer_dense(units = 3, activation = 'softmax')
model %>% compile(
     loss = 'sparse_categorical_crossentropy',
     optimizer = 'adam',
     metrics = 'accuracy'
     )
# Fit the model 
model %>% fit(
     iris.training, 
     iris.trainLabels, 
     epochs = 200, 
     batch_size = 5, 
     validation_split = 0.2
 )
```

History visualisation of model

```{r}
history <- model %>% fit(
     iris.training, 
     iris.trainLabels, 
     epochs = 200,
     batch_size = 5, 
     validation_split = 0.2
 )
plot(history)
```
Plots of loss of model

```{r}
plot(history$metrics$loss, main="Model Loss", xlab = "epoch", ylab="loss", col="blue", type="l")
lines(history$metrics$val_acc, col="green")
legend("bottomright", c("train","test"), col=c("blue", "green"), lty=c(1,1))
```
Plots of accuracy of model

```{r}
 
plot(history$metrics$acc, main="Model Accuracy", xlab = "epoch", ylab="accuracy", col="blue", type="l")
lines(history$metrics$val_acc, col="green")
legend("bottomright", c("train","test"), col=c("blue", "green"), lty=c(1,1))

```
Label Prediction

```{r}
classes <- model %>% predict_classes(iris.test, batch_size = 128)
table(iris.testtarget, classes)
```

Model evaluation

```{r}
score <- model %>% evaluate(iris.test, iris.testLabels, batch_size = 128)
print(score)

```

Adding layers to increase model fineness

```{r}
model <- keras_model_sequential() 

# Add layers to model
model %>% 
    layer_dense(units = 8, activation = 'relu', input_shape = c(4)) %>% 
    layer_dense(units = 5, activation = 'relu') %>% 
    layer_dense(units = 3, activation = 'softmax')


model %>% compile(
     loss = 'sparse_categorical_crossentropy',
     optimizer = 'adam',
     metrics = 'accuracy'
 )


model %>% fit(
     iris.training, iris.trainLabels, 
     epochs = 200, batch_size = 5, 
     validation_split = 0.2
 )

# Evaluate the model
score <- model %>% evaluate(iris.test, iris.testLabels, batch_size = 128)

# Print the score
print(score)

```

